---
title: "Market Basket Analysis in R"
author: "Oloyede Abdulganiyu"
date: "2023-04-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What kind of analysis is Market Basket Analysis again?

## Behind the scenes

*NB: you can skip this if you're not interested in learning the technologies and how it works*

As you already know or not, MBA is an unsupervised machine learning methodology, in the sense that it does not have labels to train it\*

In a typical MBA data format, the individual rows are transactions while the columns are items

\*\* insert pics of data format

Before we go deeper into MBA, we'd like to explain some terms and concepts in MBA and they are:

-   **Density**:

-   **Antecedent:** Also called the LHS (i.e: Left Hand Side). This is the item ...\* for example ...

-   **Consequent:** Also called the RHS (i.e: Right Hand Side). This is the item we're interested in

    $X => Y$ is a rule where X is the antecedent and

    Y is the consequent

-   **Support**: This is the relative frequency of the item(s) of interest in relation with the total frequency, we can as well say it is the probability of item(s) of interest. $supp(X) = prob(X) = freq(X)/n$ ; where n = total frequency.

    $supp(XY) = prob(XY) = freq(XY)/n$ ; where freq(XY) counts the number of times X and Y appear together.

-   **Confidence:** This is a measure of the reliability of the rule. Confidence is the conditional probability of the consequent beign bought together withthe antecedent. For example:

    in terms of probability ------ $conf(X,Y) = prob(Y|X) = prob(XY)/prob(X)$

    and in terems of support ---- $conf(X,Y) = supp(X,Y)/ supp(X)$

    A confidence of 0.5 means 50% of X purchases include Y purchases too. This will enable the analyst know which products pair together and how frequently they do.

-   **Lift:** Lift is the ratio of observed support to the expected support. In lay man terms, lift is the ratio of likelihood that an antecedent is bought with a consequent (item of interest) compared to any other item.

    We are most interested in lift \> 1

    Mathematical definition of **lift**:

    in terms of probability ....... $lift(X,Y) = prob(YX)/prob(X).prob(Y)$

    in terms of support ........... $lift(X,Y) = supp(X.Y)/ supp(X).supp(Y)$

    Its okay if you don't quite grab it yet, let's see an example eal quick.

    for rule $X => Y$ , the $lift(X,Y) = 1.42$ The interpretation of this lift will be: customers who bought X are 42% more likely to buy Y than any other randomly chosen item in the superstore.

    One of the interesting features of lift is that lift is symmetrical. lift(X,Y) = lift(Y,X) = 1.42

-   

-   \-

-   \-

## The Process

The project was started by installing the neceesary packages like

-   `arules` - for the apriori algorithm

-   `arulesViz` -for visualizing the affinity analysis

-   `readxl` - since the dataset is saved as csv file, we'll need to import it into R using a library, and this is that library.

-   `tidyverse` - for transformation and data cleaning

-   `plyr` and `dplyr` - for transforming the dataset into a format that will be understood by `arules`

-   `magrittr` -

    installed the packages using the `install.packages()` function and called them using the `library()` function.

Data importation was done using the code chunk :

```{r eval=FALSE, echo = TRUE}
library("readxl")
df<- read_excel(file.choose())
```

After importing the online retail store data gotten from <link>, the data cleaning process was quick as the data was not like rolling in mud, 🤣 lolz 😅.

The data transformation was more than quite tough as the data was in a format that would not be easily understood by the R libraries necessary: arules and arulesViz the data was transformed to three (3) different formats:

-   one hot encoded format

-   list separated by comma format

-   class transaction format (wide)

-   (long)

Running association rules

explore generated rules

filter generated rules

MBA visualization When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
